End-User Software Customization by Direct Manipulation of Tabular Data

### 2020-08-15 paper.tex
done

A helpful future addition would be to allow end users to create wrappers
themselves, so that more people could contribute to the shared library.
There are many existing projects exploring end-user web scraping (such
as Helena \citep{chasins2018}) which might prove helpful; however,
because the needs of a Wildcard adapter are slightly more complex than a
traditional web scraper (as discussed in
Section~\ref{sec:architecture}), further work might be needed to enable
end-user wrapper creation.

The distribution mechanism for semantic wrappers is also important for
encouraging an ecosystem of shared wrappers. Currently, the distribution
mechanism is simply merging the code for all adapters into the main
Wildcard codebase. This is the simplest workable solution, but has its
downsides: contributing involves a fairly high barrier of creating a
pull request on Github, and using newly contributed wrappers requires
installing a new version of the extension. In the future we might
explore other mechanisms, like an online repository that the extension
pulls from dynamically. Security is also a consideration---DOM scraping
adapters can execute arbitrary Javascript code, which opens up the
possibility of malicious adapters being contributed. Currently we plan
to solve this challenge with centralized code review, but another
approach we are considering is using or inventing a more restricted
domain-specific language for specifying scraping logic.

\hypertarget{generic-data-interfaces}{%
\subsection{Generic data interfaces}\label{generic-data-interfaces}}

\begin{itemize}
\item
  what makes a UI good for customization?

  \begin{itemize}
  \tightlist
  \item
    generic: can show data from many apps
  \item
    neutral: doesn't prioritize a particular use case, leaves the
    imagination open
  \item
    good affordances for performing common actions
  \end{itemize}
\item
  could be others besides a table

  \begin{itemize}
  \tightlist
  \item
    tree?
  \item
    text box?
  \item
    compose these?
  \end{itemize}
\end{itemize}

>>>>

The distribution mechanism for semantic wrappers is also important for
encouraging an ecosystem of shared wrappers. Currently, the distribution
mechanism is simply merging the code for all adapters into the main
Wildcard codebase. This is a simple solution, but makes it fairly
difficult to contribute new wrappers and requires installing a new
version of the extension to gain access to new warppers. In the future
we might explore other mechanisms, like an online repository that the
extension pulls from dynamically. Security is also a consideration---DOM
scraping adapters can execute arbitrary Javascript code, which means a
malicious adapter could exfiltrate sensitive information from a page.
Approaches to security could include centralized code review, using a
restricted scraping DSL, or creating a sandboxed context for scrapers
without access to networking APIs.

\hypertarget{alternate-mechanisms-for-wrapper-creation}{%
\subsubsection{Alternate mechanisms for wrapper
creation}\label{alternate-mechanisms-for-wrapper-creation}}

Requiring programming to create wrappers has an obvious limitation. If
an end user wants to customize a site and no programmer has contributed
a wrapper for that site, then they have no means of customizing it.
Although a sufficiently vibrant community of programmers could produce
wrappers for many popular sites, it's unrealistic to imagine covering
all websites that end users might want to customize. We envision two
strategies for dealing with this problem:

\textbf{End user wrapper creation:} If end users could create wrappers
without programming, there would be fewer limitations on the sites they
could customize. There are existing projects exploring end-user web
scraping (such as Helena \citep{chasins2018}) which might prove helpful;
however, because the needs of a Wildcard adapter are slightly more
complex than a traditional web scraper (as discussed in
Section~\ref{sec:architecture}), further work might be needed to enable
end-user wrapper creation.

\textbf{First party wrapper creation:} An integrated adapter installed
by the developer of an application could directly access internal state,
providing the same functionality as a DOM scraping adapter but in a more
robust way.

With the advent of rich frontend web frameworks, structured application
state is now often available in the web client. We suspect it is
possible to create plugins for frontend frameworks that expose this
state to Wildcard with only minimal effort from the application
developers. Customizability is sometimes a key selling point for
software. An integrated website adapter would provide a way for
developers to integrate with an ecosystem of formulas and customization
tools without needing to build all that functionality from scratch.
